\section{Background}
\label{sec:background}

vFree ~\cite{vFree} is composed of two primary strategies which, when combined, make it possible to achieve near-native performance for network-heavy workloads while still maintaining the security benefits of memory protection provided by the vIOMMU. This is done by rewriting the invalidation routines to be more efficient (and configurable), improving support for contiguous page frees, and allowing for deferred free page (DFP) which allows any drivers that use the feature to register a callback that is invoked when a page is unmapped. This allows for the driver to work on other tasks while the unmapping is being processed, which greatly reduces locking overhead and still allows for safety as the driver can release the page when it is safe to do so. The vFree ~\cite{vFree} strategy has been shown to significantly reduce the performance overhead of the vIOMMU, particularly in network-heavy workloads. However, most of the existing research on vFree ~\cite{vFree} has focused on single VM workloads, and there is a lack of comprehensive evaluation of these strategies in multi-VM environments. This paper aims to fill this gap by evaluating the performance of vFree ~\cite{vFree} in a multi-VM environment and comparing it to traditional approaches for managing the vIOMMU. By doing so, we can gain insights into the effectiveness of vFree ~\cite{vFree} in real-world scenarios where multiple VMs are running concurrently and sharing resources.

\subsection{Machines}
\label{subsec:back_bench}
In our experiment setup there are two servers called Nexus 1 and Nexus 3; Nexus 1 is where we run VMs and gather metrics on those VMs. A transmit (TX) path experiment will setup Nexus 1 as a client and We use two servers directly connected to each other via 400Gbps Nvidia ConnectX-7 NICs to ensure performance bottlenecks not in the network. The servers use Emerald Rapids processor architecture that supports nested translation, with Intel Xeon Platinum 8592V CPUs and Intel Xeon Gold 6538Y+ CPUs respectively. The former hosts our VMs, and the latter acts as client to send or receive traffic to the VM. The host server, guest VM, and client machines run with Ubuntu 22.04 and Linux v6.12.9. The VMs are hosted by QEMU v10.0 with nested translation ~\cite{yiliu1765qemu}. We place the VMs on the NIC-local NUMA node. Every vCPU of the VM is pinned to a separate physical core on the host to avoid interference. 

\subsection{Measurements}
\label{subsec:measurements}
We use Iperf3 ~\cite{iperf3} to generate network traffic with 4K MTU size, and measure throughput. We run each experiment five times, and present the average results and standard deviation. We use two setups to isolate the impact of IO memory protection on Rx and Tx datapaths. In the Rx-server setup, the client sends large payloads to the VM; the VM receives these payloads and transmits TCP ACKs back to the client. In the Tx-server setup, the VM sends large payloads to the client and receives TCP ACKs from the client.

The rest of this paper focuses on our machine setup to benchmark the three previously described strategies. Section \ref{sec:methodology} describes the methods used to obtain our results focusing mostly on the machine configuration. Section \ref{sec:results} discusses the results obtained from our experiments. 


